Collision rates of planetesimals near mean-motion resonances

### 2020-12-08 refereeResponse.txt

We appreciate the referee's thoughtfulness and perspective and would firstly like to thank them for care and thoroughness that was put into their reviewing of this paper. We will address comments from the referee one by one, with the original comment listed first, followed by our response (indented).

1) It is not clear how collisions were treated. Wallace et al. point to other papers that describe the implementations of the collisional algorithms. Although I agree they should not attempt to describe all details of the code here, it would be of major help to add more text describing how they deal with collisions.

1.1) What is the collisional scaling used?

	All collisions result in perferct mergers. This is now stated in section 3.1. Because we only model where the collisions happen, rather than what the debris looks like, we keep the collision model simple.

1.2) What material is assumed for the colliding planetesimals?

	A 2 g cm^-3 bulk density corresponds to an even mixture of ice and rock. Because all collisions are assumed to result in perfect mergers, this only affects the mass and radius of the bodies.

1.3) How do you deal with size frequency distribution (collisional cascade) resultant from collisions?

	We do not model the evolution of the resulting debris from collisions. More on this below.

1.4) What is the dust size considered in this work?

	We assume that the resulting dust grains are 1mm in size. This is mentioned in section 4.2.

@@

I would say all these matter a lot once 300 km planetesimals will likely not be instantaneously pulverized to dust-size when they collide. Thus it is of major importance for the understanding of this work, to know precisely how collisions were treated. If, for instance, authors did not consider any fragmentation, and only tracked collisions to determine where dust would form, I would say this should be put as a caveat.

The reason why I'm asking this to be a caveat is because in this case the authors' results would not necessarily be representing the dust profile, but rather the planetesimal collisional profile, and so the upper limit of dust generation at a given location. This makes a difference regarding the applicability of this work, once the expected real signals from the bumps and dips would be much smaller, and so particularly undetectable. The conclusion's by Wallace et al., potentially based on a local upper limit of dust generation, already points towards an expected faint signal in real observation. Therefore, if the real expected signal end up being even fainter, would this work results be applicable at all, or would them be simply negative results? I am afraid that if the results would be simply negative, they may not be important. I mean, why should we care if we will not be able to easily see those bump/dip features? In this case we would not be able to say anything about the mass and eccentricity of the potential perturbing planet. Therefore, what is the meaning of constraining something that it is unmeasurable?

I would try to help the authors and turn this into a positive light, even if the results turn to be negative. One could eventually say that, once observational capabilities improve, having a constrained measurement would help to quickly interpret images in the future. I don't know how likely is this, but with a proper caveat, it would be acceptable, eventually.

	This is an excellent point brought up by the referee and is something that deserves a more careful discussion of in the paper. In regards to the applicability of this work: one must first understand where and how the planetsimals collide if there is any hope of nailing down what the resulting dust distribution will look like. Going from the collision profile to a dust emission profile deserves an entire paper by itself and we are certainly not doing that analysis justice here.

	In light of this, we think that the best thing to do is to reframe this paper a bit, so that the collisions->dust step is saved until the end, given a little less weight in the abstract, introduction and conclusion, and the concerns brought up by the referee are sufficiently emphasized in the text. Most importantly, the reader should be cautioned that the dust profiles we present are upper limits.

	Both the abstract and the summary at the end of the introduction have been reworked to place more emphasis on the planetesimal dynamics and collisions, with the dust profiles and giant planet property constraints as more of an afterthought. Sections 2, 3 and 4 of the paper now focus solely on the 300 km planetesimals, save for a few places where we justify our neglect of gas drag during the simulations. All discussion of dust production and its observability is now saved for the latter parts of the paper.

@@

2) I am very concerned by the fact the authors neglected migration on small particles due to aerodynamic drag. 300 km planetesimals will not break down to dust directly (of course some dust will be created during the 300 km planetesimal's collisions, but that would be a very small fraction of mass). The 300 km planetesimals (depending on material) will more likely break into a collection of smaller km-size objects. These smaller size objects will then overpopulate the regions where they were created, thus enhancing collision rates among them. Collisions that continue to happen among these smaller objets will eventually turn them in to a very large swarm of even smaller m- to km-size objects. Ultimately, this last generation of collisionally evolved planetesimals will become dust (then my question about how and if Wallace et al. considered collisional cascade in their simulations).

The problem of neglecting aerodynamic drag becomes then a bigger problem because, as it is very well known, small m- to km-sized objects feel strong headwind and thus can drift very fast Sunwards (e.g. Weidenschilling & Davis 1985; Patterson 1987; Beauge & Ferraz-Mello 1993; Malhotra 1993; Gomes 1995a,b; Jiang & Yeh 2004; Batigyn & Laughlin 2015; Storch & Batygin 2019; Deienno et al. 2020). Therefore, dust created by these large moving swarms could leave a signature that is far different than that when assuming dust would be generated in the location of the 300 km planetesimal collisions (eventually changing, or even destroying, the bump/dip signatures; despite the fact that dust would couple to the gas almost instantaneously, once created). This is why I previously said authors may be tracking the planetesimal collisional profile and the upper limit of dust generation at those locations in their simulations, but not necessarily the dust profile.

	In the original paper, we checked to ensure that the 300 km planetsimals and the < mm-sized dust grains would stay in place, but the referee's concern about the effects of radial drift of intermediate-sized bodies is an important point that was overlooked. We handle this by: 1. cautioning the reader that the dust emission profiles presented should be interpreted as upper limits 2. discussing the concerns brought up by the referee 3. Providing a (somewhat crude) calculation of the drift and collision timescale of 1m bodies, where drift should be maximal. Point 3 is discussed more below.

@@

I would suggest that the authors investigate and say explicitly in the text what is the expected relation between collisional timescale and drift timescale for m- to km-sized (1 m - 1 km) particles.

They should discuss:
2.1) If drift timescale of 1 m- to 1 km-sized objects is larger than their collision timescale.
In this case, collisions would happen before planetesimals would drift. Then the results should hold true.
2.2) If drift timescale of 1 m- to 1 km-sized objects is smaller than their collision timescale.
In this case, planetesimals would tend to drift faster than they can collisionally evolve. However, if their drift timescale is still larger than 5000 yr (integration time), then maybe this could be a minor issue. But a caveat would still be necessary as the large swarm of small planetesimals would likely continue collisionally evolving elsewhere (far from the mean motion resonances; MMRs, thus potentially changing the bump/dip signatures).
2.3) If drift timescale is smaller than 5000 yr and also smaller than collisional timescale.
In this case, this is a major issue! Planetesimals would quickly fall into the star, leaving minor dust signature, or a continue dust signal throughout their migration path, thus certainly no bumps nor dips would be observed.

Regardless the situation, this should be clearly discussed in the text.

	In section 5, we now provide a calculation of both the drift timescale and collision timescale for 1m bodies near the 2:1 MMR. Because we do not model the debris from planetesimal collisions, the collision timescale for the 1m bodies is found by extrapolating the collision rate of the planetesimals to smaller bodies. This is done in the following way:

		1. Measure the collision rate (per particle) of the planetesimals from the e2m2 simulation inside the 2:1 MMR
		2. Assume that the 1m bodies will have dynamical state (eccentricity and inclination dispersion) as the planetesimals
		3. Assume that all of the mass in solids near the 2:1 MMR is now made up of 1m bodies
		4. Scale the measured planetesimal collision rate by the change in number density and collision cross section (if one were to convert the planetesimal population to smaller bodies) to get the collision timescale for 1m bodies.

	In doing so, we find that the collision timescale for 1m bodies is about 100x smaller than the drift timescale for those bodies across the resonance. This suggests that radial drift will not have a large influence on the resulting dust dustribution. However, the 1m collision timescale should be interpreted as an lower limit, due to assumptions 2. and 3. We conclude that it is therefore plausible that radial drift does not play an important role, although a proper treatment of the collisional cascade and drift process is necessary to actually confirm this.

### 2020-12-08 main.tex
More edits, responses to referee in report

The dynamics governing the motion of bodies near MMRs is extremely nonlinear, as is determining what the collision rates between planetesimals 
should look like in these regions. For a collection of bodies massive enough to experience the effects of gravitational focusing, a large eccentricity 
dispersion tends to reduce the probability of collision, while enhancements in surface density tends to increase it. Due to conservation of the Jacobi 
energy, MMRs simultaneously enhance the local eccentricity dispersion and also enhance the surface density adjacent to the resonance 
\citep{2000Icar..143...45R, 2017ApJ...850..103B}. Unfortunately, collision detection in an N-body simulation is extremely computationally expensive. 
So far, studies of planetesimal dynamics near MMRs \textbf{in which the planetesimals are directly resovled} have involved either collisionless test particles 
\citep{2017ApJ...850..103B, 2016ApJ...818..159T, 2018ApJ...857....3T} or severely limited integration times \citep{2000Icar..143...45R}.

@@

\textbf{\noindent where $m_{pl}$ and $s_{pl}$ are the mass and radius of the planetesimal. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. At 3 au, the gaseous component of the solar nebula has a density of $4 
\times 10^{-10}$ g cm$^{-3}$ and the typical headwind experienced by a body on a Keplerian orbit is $\sim 5000$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 100 km body with a 
density of 2 g cm$^{-3}$ and an eccentricity of 0.1, the stopping timescale is on the order of $10^{6}$ years. This is much longer than the timescales related to secular forcing and libration near mean-motion resonances, as discussed above, but shorter than the typical protoplanetary disk lifetime and potentially the planetesimal formation timescale. For this reason, we do not model the effects of gas drag, although we consider its effects when constructing initial conditions for our simulations, which is discussed next.}

@@

To follow the dynamical and collisional evolution of a planetesimal disc, we use the highly parallel N-body code {\sc ChaNGa} 
\footnote{A public version of {\sc ChaNGa} can be downloaded from \url{http://www-hpcc.astro.washington.edu/tools/ChaNGa.html}}. 
This code, which is written in the {\sc CHARM++} parallel programming language, was originally designed for cosmology simulations 
and has been shown to perform well on up to half a million processors \citep{2015AphCom..2..1}. \textbf{Using similar methods to {\sc PKDGRAV}, which has been used for numerous studies of planet formation (add some references here)}, {\sc ChaNGa} calculates 
gravitational forces using a modified Barnes-Hut \citep{1986Natur.324..446B} tree with hexadecapole expansions of the moments \textbf{and then integrates the equations of motion using a kick-drift-kick leapfrog scheme}. 
All of the simulations we perform use a node opening criterion of $\theta_{BH}$ = 0.7. More information about the code can be found 
in \citet{2008IEEEpds...ChaNGa}.

@@

\noindent with $\alpha$ = 3/2 and $\Sigma_{0}$ = 10 g cm$^{-2}$. Planetesimals are given a bulk density of 2 g cm$^{-3}$ and a diameter of 
300 km. \textbf{This choice of parameters for the planetesimal disk is meant to follow the setup used in simulation B by \citet{2000Icar..143...45R}. One difference from the aformentioned study is that we use a narrower annulus, which reduces the number of particles required from $10^6$ down to roughly 500,000. This allows us to use a finer base timestep size, which, as we will discuss in section \ref{sec:results}, appears to be the reason why no features were seen in the collision profile near the resonances by \citet{2000Icar..143...45R}}. Each simulation is evolved for 5,000 years, which is about 400 orbits 
of the perturbing planet.

@@

\subsection{Where Does the Dust End Up?}\label{sec:dust}

>>>>

The dynamics governing the motion of bodies near MMRs is extremely nonlinear, as is determining what the collision rates between planetesimals 
should look like in these regions. For a collection of bodies massive enough to experience the effects of gravitational focusing, a large eccentricity 
dispersion tends to reduce the probability of collision, while enhancements in surface density tends to increase it. Due to conservation of the Jacobi 
energy, MMRs simultaneously enhance the local eccentricity dispersion and also enhance the surface density adjacent to the resonance 
\citep{2000Icar..143...45R, 2017ApJ...850..103B}. Unfortunately, collision detection in an N-body simulation is extremely computationally expensive. 
So far, studies of planetesimal dynamics near MMRs \textbf{in which the planetesimals are directly resovled} have involved either collisionless test particles 
\citep{2017ApJ...850..103B, 2016ApJ...818..159T, 2018ApJ...857....3T} or \textbf{integration times that do not fully capture the dynamisc of the resonances} \citep{2000Icar..143...45R, 2013ApJ...777L..31D}.

@@

\textbf{\noindent where $m_{pl}$ and $s_{pl}$ are the mass and radius of the planetesimal. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. At 3 au, the gaseous component of the solar nebula has a density of $4 
\times 10^{-10}$ g cm$^{-3}$ and the typical headwind experienced by a body on a Keplerian orbit is $\sim 5,000$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 100 km body with a 
density of 2 g cm$^{-3}$ \textbf{(assuming a mixture of ice and rock)} and an eccentricity of 0.1, the stopping timescale is on the order of $10^{6}$ years. This is much longer than the timescales related to secular forcing and libration near mean-motion resonances, as discussed above, but shorter than the typical protoplanetary disk lifetime and potentially the planetesimal formation timescale. For this reason, we do not model the effects of gas drag, although we consider its effects when constructing initial conditions for our simulations, which is discussed next.}

@@

To follow the dynamical and collisional evolution of a planetesimal disc, we use the highly parallel N-body code {\sc ChaNGa} 
\footnote{A public version of {\sc ChaNGa} can be downloaded from \url{http://www-hpcc.astro.washington.edu/tools/ChaNGa.html}}. 
This code, which is written in the {\sc CHARM++} parallel programming language, was originally designed for cosmology simulations 
and has been shown to perform well on up to half a million processors \citep{2015AphCom..2..1}. \textbf{Using similar methods to {\sc PKDGRAV}, which has been used for numerous studies of planet formation \citep{2000Icar..143...45R, 2005ApJ...625..427L, 2013ApJ...777L..31D, 2015ApJ...806...23L}}, {\sc ChaNGa} calculates 
gravitational forces using a modified Barnes-Hut \citep{1986Natur.324..446B} tree with hexadecapole expansions of the moments \textbf{and then integrates the equations of motion using a kick-drift-kick leapfrog scheme}. 
All of the simulations we perform use a node opening criterion of $\theta_{BH}$ = 0.7. More information about the code can be found 
in \citet{2008IEEEpds...ChaNGa}.

@@

\noindent with $\alpha$ = 3/2 and $\Sigma_{0}$ = 10 g cm$^{-2}$. Planetesimals are given a bulk density of 2 g cm$^{-3}$ \textbf{(assume a mixture of ice and rock)} and a diameter of 
300 km. \textbf{This choice of parameters for the planetesimal disk is meant to follow the setup used in simulation B by \citet{2000Icar..143...45R}. One difference from the aformentioned study is that we use a narrower annulus, which reduces the number of particles required from $10^6$ down to roughly 500,000. This allows us to use a finer base timestep size, which, as we will discuss in section \ref{sec:results}, appears to be the reason why no features were seen in the collision profile near the resonances by \citet{2000Icar..143...45R}}. Each simulation is evolved for 5,000 years, which is about 400 orbits 
of the perturbing planet.

@@

\section{Where Does the Dust End Up?}\label{sec:dust}
